\begin{otherlanguage}{english}
    \begin{abstract}
    \thispagestyle{empty}
This work investigates the feasibility of using facial recognition based on \textit{embeddings}, vector representations of facial characteristics, as an alternative to \textit{Near Field Communication} (NFC) cards used for user identification in institutional environments. Initially, a conceptual and experimental validation of the approach was conducted using pretrained models for facial feature extraction and similarity metrics in the vector space. This stage employed a synthetic facial image dataset, adopted in compliance with ethical considerations and the Brazilian General Data Protection Law (LGPD), enabling performance evaluation through the biometric metrics False Acceptance Rate (FAR) and False Rejection Rate (FRR).

The experimental results demonstrated clear separation between different identities and consistency among samples of the same identity, indicating the technical feasibility of the proposed solution. Based on these findings, a functional real time facial recognition prototype was developed using a webcam, local biometric data storage, and a graphical user interface for user enrollment, management, and identification. The system employs the InsightFace library through the \textit{FaceAnalysis} module and the \textit{buffalo\_l} model package for face detection, landmark localization, facial alignment, and ArcFace embedding generation. In addition, multipose enrollment, capture quality assessment, and temporal voting mechanisms were implemented to improve recognition robustness.

The final results show that the proposed approach is technically viable as an alternative to physical cards, combining experimental evidence with practical validation through a functional prototype capable of performing automated and contactless biometric identification.
\\[0.5em]
\textbf{Keywords:} facial recognition; biometrics; \textit{embeddings}; computer vision; access control; \acs{NFC}.
\end{abstract}
\end{otherlanguage}
